---
title: "Bellabeat_analysis"
output: html_notebook
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

### import libraries
```{r}
#install.packages('tidyverse')
library(tidyverse) 
#install.packages('foreign')
library(foreign)
#install.packages('ggplot2')
library(ggplot2)
library("scales")
#install.packages('plyr')
#library(plyr)
```

## Load datasets
#### fitbit data & garmin data & sleep data
```{r}
dailyActivity_fitbit <- read.csv("/Users/jordynohashi/Desktop/DA/bellabeatCS/FitabaseData/dailyActivity_merged.csv")
data_study <- read.csv("/Users/jordynohashi/Desktop/DA/bellabeatCS/dataverse_files/data_raw.csv",sep=",")
sleep_data <- read.xport("/Users/jordynohashi/Desktop/DA/bellabeatCS/P_SLQ.xpt")
```

### Clean up data
#### sleep data rename cols
```{r}
sleep_data <- sleep_data %>% 
  dplyr::rename('weekdayHRS'='SLD012', 'weekendHRS'='SLD013', 'snore'='SLQ030', 'bad_breathing'='SLQ040', 'bad_sleep'='SLQ050', 'sleepy'='SLQ120') 
```

#### remove duplicate data
```{r}
dailyActivity_fitbit <- dailyActivity_fitbit %>% distinct()
data_study <- data_study %>% distinct()
sleep_data <- sleep_data %>% distinct()
```

#### Remove missing vals
```{r}
sleep_data <- sleep_data %>% drop_na(weekdayHRS,weekendHRS)
```

### Group sleep hours by categories 
less than 6: Not recommended sleep, 1
6: Appropriate sleep, 2
7-9: Recommended sleep, 3
10-11: Appropriate sleep, 2 
more than 11: Over sleep, 4
```{r}
# sleep label
sleep_data <- sleep_data %>% mutate(Sleep =
  case_when(weekdayHRS < 6 ~ "Not recommended sleep",
            weekdayHRS == 6 | weekdayHRS < 7 ~ "Appropriate sleep",
            weekdayHRS >= 7 & weekdayHRS <= 9 ~ "Recommended sleep",
            weekdayHRS > 9 & weekdayHRS <= 11 ~ "Appropriate sleep",
            weekdayHRS > 11 ~ "Over sleep")
)

# sleep number
sleep_data <- sleep_data %>% mutate(SleepNum =
  case_when(weekdayHRS < 6 ~ 1,
            weekdayHRS == 6 | weekdayHRS < 7 ~ 2,
            weekdayHRS >= 7 & weekdayHRS <= 9 ~ 3,
            weekdayHRS > 9 & weekdayHRS <= 11 ~ 2,
            weekdayHRS > 11 ~ 4)
)
```

### Merge two datasets -> df
columns: id, date, steps, calories, sedMin, lightMin, modMin, veryMin, provider

#### data_study -> df_study
convert date, ID->id, Date->date, Steps->steps, TEE->calories, Sedentary->sedMin, 
LPA->lightMin, MPA->modMin, VPA->veryMin, Provider->provider
```{r}
data_study$date <- strptime(as.character(data_study$Date), "%d.%m.%Y")
data_study$date <- format(data_study$date, "%m/%d/%Y")
df_study <- data_study[, c("ID","date","Steps","TEE","Sedentary","LPA","MPA","VPA","Provider")]
df_study <- df_study %>% 
  dplyr::rename('id'='ID', 'steps'='Steps', 'calories'='TEE', 'sedMin'='Sedentary', 
         'lightMin'='LPA', 'modMin'='MPA', 'veryMin'='VPA', 'provider'='Provider')
```

#### dailyActivity_fitbit -> fitbit_study
Id->id, ActivityDate->date, TotalSteps->steps, VeryActiveMinutes->veryMin,
FairlyActiveMinutes->modMin, LightlyActiveMinutes->lightMin, Calories->calories, add provider
```{r}
fitbit_study <- dailyActivity_fitbit[, c("Id","ActivityDate","TotalSteps","Calories", "SedentaryMinutes", "LightlyActiveMinutes", "FairlyActiveMinutes", "VeryActiveMinutes")]
fitbit_study <- fitbit_study %>% 
  dplyr::rename('id'='Id', 'date'='ActivityDate', 'steps'='TotalSteps','calories'='Calories', 'sedMin'='SedentaryMinutes', 'lightMin'='LightlyActiveMinutes',
         'modMin'='FairlyActiveMinutes', 'veryMin'='VeryActiveMinutes',)
fitbit_study$provider <- "Fitbit"
```

```{r}
df <- rbind(df_study, fitbit_study)
glimpse(df)
```

### filter out valid days(>= 150 steps)
```{r}
df_val <- df[df$steps>=150,]
glimpse(df_val)
```

#### convert date character to date type
```{r}
typeof(df_val$date)
df_val$date <- as.Date(df_val$date, "%m/%d/%Y")
df_val$num <- as.numeric(df_val$date, "%m/%d/%Y")
typeof(df_val$date)
```

### create new data frame with id, number of observations, average steps, calories, activities, provider
```{r}
df_summary <-
  df_val %>%
  group_by(id) %>%
  mutate(count = n()) %>%
  summarise(avg_steps=mean(steps),
            avg_calories=mean(calories),
            avg_sedMin=mean(sedMin),
            avg_lightMin=mean(lightMin),
            avg_modMin=mean(modMin),
            avg_veryMin=mean(veryMin),
            provider=first(provider),
            count=mean(count))
```



### Check the summary
```{r}
summary(df_val)
```

```{r}
summary(df_summary)
```

```{r}
df_fitbit <- df_val[df_val$provider=="Fitbit",]
summary(df_fitbit)
```

```{r}
df_garmin <- df_val[df_val$provider=="Garmin",]
summary(df_garmin)
```

```{r}
df_sum_large <- df_summary[df_summary$count>100,]
df_sum_fitbit <- df_sum_large[df_sum_large$provider=="Fitbit",]
summary(df_sum_fitbit)
```

```{r}
df_sum_garmin <- df_sum_large[df_sum_large$provider=="Garmin",]
summary(df_sum_garmin)
```


```{r}
summary(sleep_data)
```


### Garmin vs Fitbit
```{r}
glimpse(df_fitbit)
glimpse(df_garmin)
```

### Steps
```{r}
ggplot(data=df_val) +  
  geom_histogram( aes(steps) ) 
```

### Calories
```{r}
ggplot(data=df_val) +  
  geom_histogram( aes(calories) ) 
```

### Sedentary physical activity
```{r}
ggplot(data=df_val) +  
  geom_histogram( aes(sedMin) ) 
```

### Light physical activity
```{r}
ggplot(data=df_val) +  
  geom_histogram( aes(lightMin) ) 
```

### Moderate physical activity
```{r}
ggplot(data=df_val) +  
  geom_histogram( aes(modMin) ) 
```

### Vigorous physical activity
```{r}
ggplot(data=df_val) +  
  geom_histogram( aes(veryMin) ) 
```

### Garmin vs Fitbit steps & calories
```{r}
ggplot(df_val, aes(steps, calories)) + 
  geom_point() +
  facet_grid( provider ~ . ) +
  labs(title = "Steps vs Calories", x='Steps', y='Calories')
```

### Garmin vs Fitbit minutes of moderate physical activity & calories
```{r}
ggplot(df_val, aes(modMin, calories)) + 
  geom_point() +
  facet_grid( provider ~ . )
```

### Garmin vs Fitbit minutes of vigorous physical activity & calories
```{r}
ggplot(df_val, aes(veryMin, calories)) + 
  geom_point() +
  facet_grid( provider ~ . ) +
  labs(title = "Vigorous physical activity vs Calories", x='Minutes of vigorous physical activity', y='Calories')
```

### Garmin vs Fitbit minutes of moderate physical activity & steps
```{r}
ggplot(df_val, aes(modMin, steps)) + 
  geom_point() +
  facet_grid( provider ~ . )
```

### Garmin vs Fitbit minutes of vigorous physical activity & steps
```{r}
ggplot(df_val, aes(veryMin, steps)) + 
  geom_point() +
  facet_grid( provider ~ . ) +
  labs(title = "Vigorous physical activity vs Steps", x='Minutes of vigorous physical activity', y='Steps')
```

### graph number of observations vs average steps, calories, activities (see common occurence and consistency effects) (also compare fitbit vs garmin)
```{r}
ggplot(data=df_summary) +  
  geom_histogram( aes(avg_steps) ) 
ggplot(data=df_summary) +  
  geom_histogram( aes(avg_calories) ) 
ggplot(data=df_summary) +  
  geom_histogram( aes(avg_sedMin) ) 
ggplot(data=df_summary) +  
  geom_histogram( aes(avg_lightMin) ) 
ggplot(data=df_summary) +  
  geom_histogram( aes(avg_modMin) ) 
ggplot(data=df_summary) +  
  geom_histogram( aes(avg_veryMin) ) 
```
### Graphs above: count represents people, shows number of people with varying avgs of steps, calories, etc.
```{r}
ggplot(df_summary, aes(count, avg_steps)) + 
  geom_point() +
  facet_grid( provider ~ . )
ggplot(df_summary, aes(count, avg_calories)) + 
  geom_point() +
  facet_grid( provider ~ . )
ggplot(df_summary, aes(count, avg_sedMin)) + 
  geom_point() +
  facet_grid( provider ~ . )
ggplot(df_summary, aes(count, avg_lightMin)) + 
  geom_point() +
  facet_grid( provider ~ . )
ggplot(df_summary, aes(count, avg_modMin)) + 
  geom_point() +
  facet_grid( provider ~ . )
ggplot(df_summary, aes(count, avg_veryMin)) + 
  geom_point() +
  facet_grid( provider ~ . )
```
#### graphs above: count is number of observations from each person in relation to their averages --> as you can see the amount you use the device doesn't effect in increase in other variables

```{r}

ggplot(df_sum_large, aes(count, avg_steps)) + 
  geom_point() +
  facet_grid( provider ~ . )
```

### sleep graphs
```{r}
# sleep_data = ddply(sleep_data, .(snore), transform,
#                    pct_snore=(SleepNum/sum(SleepNum) * 100))
# 
# sleep_data = ddply(sleep_data, .(snore), transform, pos_snore = (cumsum(SleepNum) - 0.5 * SleepNum))
# 
# sleep_data$pct_snorelabel = paste0(sprintf("%.0f", sleep_data$pct_snore), "%")

```


```{r}
ggplot(sleep_data, aes(factor(snore),fill=Sleep ) ) +
  geom_bar(position = "stack") +
  scale_x_discrete(name = "How often do you snore?", 
                   labels=c("0" = "Never", "1" = "1-2 nights a week",
                              "2" = "3-4 nights a week",
                            "3" = "5 or more nights a week",
                            "7" = "Refused",
                            "9" = "Don't know")) +
  theme(axis.text.x = element_text(angle = 45,hjust=1)) 
```


```{r}
ggplot(sleep_data, aes(factor(bad_breathing),fill=Sleep) ) +
  geom_bar(position = "stack") +
  scale_x_discrete(name = "How often do you snort or stop breathing?", 
                   labels=c("0" = "Never", "1" = "1-2 nights a week",
                              "2" = "3-4 nights a week",
                            "3" = "5 or more nights a week",
                            "7" = "Refused",
                            "9" = "Don't know")) +
  theme(axis.text.x = element_text(angle = 45,hjust=1))
```


```{r}
ggplot(sleep_data, aes(factor(bad_sleep),fill=Sleep ) ) +
  geom_bar(position = "stack") +
  scale_x_discrete(name = "Ever told doctor had trouble sleeping?", 
                   labels=c("1" = "Yes", "2" = "No",
                            "7" = "Refused",
                            "9" = "Don't know")) +
  theme(axis.text.x = element_text(angle = 45,hjust=1))
```

```{r}
ggplot(sleep_data, aes(factor(sleepy),fill=Sleep)) +
  geom_bar(position = "stack") +
  scale_x_discrete(name = "How often feel overly sleepy during day?", 
                   labels=c("0" = "Never", "1" = "Rarely - 1 time a month	",
                              "2" = "Sometimes - 2-4 times a month",
                            "3" = "Often- 5-15 times a month",
                            "4" = "Almost always - 16-30 times a month",
                            "7" = "Refused",
                            "9" = "Don't know")) +
  theme(axis.text.x = element_text(angle = 45,hjust=1))
```


```{r}
ggplot(sleep_data, aes(factor(snore),fill=Sleep ) ) +
  geom_bar(position = "stack") +
  scale_x_discrete(name = "How often do you snore?", 
                   labels=c("0" = "Never", "1" = "1-2 nights a week",
                              "2" = "3-4 nights a week",
                            "3" = "5 or more nights a week",
                            "7" = "Refused",
                            "9" = "Don't know")) +
  theme(axis.text.x = element_text(angle = 45,hjust=1)) 
```
#### Week day sleep hours
```{r}
ggplot(sleep_data, aes(weekdayHRS) ) +
  geom_bar()
ggplot(sleep_data, aes(weekendHRS) ) +
  geom_bar()
```

#### Sleep vs attributes(not informative)
```{r}
ggplot(sleep_data, aes(weekdayHRS, snore)) + 
  geom_point()
ggplot(sleep_data, aes(weekdayHRS, bad_breathing)) + 
  geom_point()
ggplot(sleep_data, aes(weekdayHRS, bad_sleep)) + 
  geom_point()
ggplot(sleep_data, aes(weekdayHRS, sleepy)) + 
  geom_point()
```
#### Look at observations over time

```{r}
df_val_2019 <- df_val[df_val$num>as.numeric(as.Date("2019-01-01")),]
ggplot(df_val_2019) + aes(x = date) + 
  geom_bar()
```
#### Look at dates of different activity levels based on calories

##### low activity
```{r}
df_low_cal <- df_val_2019[df_val_2019$calories<2152,]
ggplot(df_low_cal) + aes(x = date) + 
  geom_bar() +
  labs(title='Low Calorie Output', x='Date', y='Count')
```

##### average activity
```{r}
df_avg_cal <- df_val_2019[df_val_2019$calories>=2152 & df_val_2019$calories<=2895 ,]
ggplot(df_avg_cal) + aes(x = date) + 
  geom_bar() +
  labs(title='Average Calorie Output', x='Date', y='Count')
```

##### high activity
```{r}
df_high_cal <- df_val_2019[df_val_2019$calories<2895 ,]
ggplot(df_high_cal) + aes(x = date) + 
  geom_bar() +
  labs(title='High Calorie Output', x='Date', y='Count')
```
##### Averages of dates
```{r}
df_low_count <-
  df_low_cal %>%
  group_by(date) %>%
  mutate(count = n()) %>%
  summarise(count=mean(count))
mean(df_low_count$count)
```

```{r}
df_avg_count <-
  df_avg_cal %>%
  group_by(date) %>%
  mutate(count = n()) %>%
  summarise(count=mean(count))
mean(df_avg_count$count)
```

```{r}
df_high_count <-
  df_high_cal %>%
  group_by(date) %>%
  mutate(count = n()) %>%
  summarise(count=mean(count))
mean(df_high_count$count)
```

### Pearson Checks
#### Compare steps & calories garmin
```{r}
cor.test(df_garmin$steps, df_garmin$calories)
```

#### Compare steps & calories fitbit
```{r}
cor.test(df_fitbit$steps, df_fitbit$calories)
```

#### Compare minutes of moderate physical activity & calories garmin
```{r}
cor.test(df_garmin$modMin, df_garmin$calories)
```

#### Compare minutes of moderate physical activity & calories fitbit
```{r}
cor.test(df_fitbit$modMin, df_fitbit$calories)
```

#### Compare minutes of vigorous physical activity & calories garmin
```{r}
cor.test(df_garmin$veryMin, df_garmin$calories)
```

#### Compare minutes of vigorous physical activity & calories fitbit
```{r}
cor.test(df_fitbit$veryMin, df_fitbit$calories)
```

#### Compare minutes of moderate physical activity & steps garmin
```{r}
cor.test(df_garmin$modMin, df_garmin$steps)
```

#### Compare minutes of moderate physical activity & steps fitbit
```{r}
cor.test(df_fitbit$modMin, df_fitbit$steps)
```

#### Compare minutes of vigorous physical activity & steps garmin
```{r}
cor.test(df_garmin$veryMin, df_garmin$steps)
```

#### Compare minutes of vigorous physical activity & steps fitbit
```{r}
cor.test(df_fitbit$veryMin, df_fitbit$steps)
```



